---
title: "CodeBook"
author: "success"
date: "2024-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, cache=TRUE}
subject.train <- readLines('subject_train.txt')
x.train <- readLines('X_train.txt')
y.train <- readLines('y_train.txt')
subject.test <- readLines('subject_test.txt')
x.test <- readLines('X_test.txt')
y.test <- readLines('y_test.txt')
```

## Introduction
In this article is describe the code and transformation I used to clean up data. The data is from collected from the accelerometer and gyroscope of the Samsung Galaxy S smartphone. The data include labels from human activities which include Sitting, Laying Down, Walking, Walking Up, Walking Down, etc. The data is presented in a *.txt* format. So before we can do any type of processing we have to load the data and format it into a data frame.

```{r code1, cache=TRUE}
library(dplyr)
data1 <- lapply(x.train, trimws) %>% as.character %>%
  strsplit('\\s+') %>% as.data.frame %>% t %>% as.data.frame

data2 <- lapply(x.test, trimws) %>% as.character %>%
  strsplit('\\s+') %>% as.data.frame %>% t %>% as.data.frame
```
The data consists of *data1* which is the training set and *data2* which is the test set.
In this study, I am only interested in the measurements on the mean and standard deviation for each meausrement. So I am going to select columns that I am interested in and merge *data1* and *data2* into a single *data* dataframe

```{r cache=FALSE}
col <- readLines('features.txt') %>% strsplit('\\s+') %>%
  lapply(function(vector){vector[2]}) %>% as.character
names(data1) <- col
names(data2) <- col
data1 <- data1[,grepl('-mean|-std', names(data1))]
data2 <- data2[,grepl('-mean|-std', names(data2))]
data <- rbind(data1, data2)
```
```{r, echo=FALSE, include=FALSE}
subject <- c(subject.train, subject.test)
activity <- c(y.train, y.test)
data$subject <- subject
data$activity <- activity
data$activity <- factor(data$activity, 
                        labels = c('WALKING', 'WALKING UPSTAIRS', 
                                   'WALKING DOWNSTAIRS', 
                                   'SITTING', 'STANDING', 
                                   'LAYING')
                        )

```
Now that I have my dataframe object, I am going to group it by *activity* and *subject*. And summarise the dataframe with he average of each variable for each activity and each subject.

```{r}
data <- lapply(data, as.numeric) %>% as.data.frame
data <- group_by(data, activity, subject) %>% 
  summarise_all(mean)
data$activity <- factor(data$activity, 
                        labels = c('WALKING', 'WALKING UPSTAIRS', 
                                   'WALKING DOWNSTAIRS', 
                                   'SITTING', 'STANDING', 
                                   'LAYING')
)
```

Now my final tidy dataset should look like this
```{r}
head(data)
```