---
title: "Plot of movie ratings against budget to see if there is any linear relationship"
author: "success(syko2021gmail.com)"
date: "2024-09-05"
output: slidy_presentation
---

# Overview
In this slide I demonstrate that movies with better ratings do not generally have higher budgets

# Libraries and Dataset
Here are the libraries and dataset I used

```{r, message=FALSE, warning=FALSE, results='hold'}
library(ggplot2movies)
library(ggplot2)
library(dplyr)
library(plotly)

data("movies")
names(movies)
```

# Data cleaning

```{r}
movies = filter(movies, !is.na(budget) & !is.na(rating))
```

# Plot
Here I make a scatter plot of movie ratings against budget and added a fitted model. The line is exactly horizontal, which suggest that there is no relationship, at least without further analysis.

```{r, echo=FALSE}
g = ggplot(movies, aes(rating,budget)) +
  geom_point() + geom_smooth(method='lm')

ggplotly(g)
```

# The code I used for the plot

```{r, eval=FALSE}
g = ggplot(movies, aes(rating,budget)) +
  geom_point() + geom_smooth(method='lm')

ggplotly(g)
```